---
title: "Figure2"
author: "Melanie Hay"
date: "2022-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(phyloseq); packageVersion("phyloseq")

# General data analysis
library(data.table); packageVersion("data.table")
library(tidyverse)
library(dplyr)
library("FactoMineR")
library("factoextra")

# For plotting
library(ggplot2); packageVersion("ggplot2")
library(scales)
library(ggpubr)
library(ggrepel)
library(scico)
library(GGally)

# set palette
ent_palette <- c("#F86B62", "#00BA38", "#619CFF")
```

# Resources  
http://www.sthda.com/english/wiki/fviz-pca-quick-principal-component-analysis-data-visualization-r-software-and-data-mining#fviz_pca_biplot-biplot-of-individuals-of-variables


## Figure 2 R Markdown

Figure 2 has three sections. We start with the PCA (A), followed by relative abundance of the enterotypes by Phylum (B) and Genus (C), and finally we have alpha diversity box plot comparisons (D, E, F, G) 


#### A) PCA

# Perform the PCA
```{r}
# First we load the clr-transformed abundance data (otu table)

PCA.381.pre <- read.csv("PCA.381.fig2.csv", row.names = 1)

# This table contains all data including the samples that fell "between" the three major enterotype clusters. Therefore, the first step is to remove the samples with a "NA" enterotype. 

PCA.381 <- filter(PCA.381.pre, Enterotype!=is.na(Enterotype))

# Principal component analysis
g.res.pca  <- PCA(PCA.381, quali.sup=382, graph = TRUE) 
#Extract and visualize eigenvalues/variances:
# Extract eigenvalues/variances
get_eig(g.res.pca)

# Visualize eigenvalues/variances
fviz_screeplot(g.res.pca, addlabels = TRUE, ylim = c(0, 30))

# Extract the results for variables
g.var <- get_pca_var(g.res.pca)

# Coordinates of variables
head(g.var$coord)

# Contribution of variables
head(g.var$contrib)

cos2.rank.pca <- data.frame(g.var$cos2)
```

# Examining the variables ( genera) 

Graphs and tables here are exploratory- not in the final manuscript
```{r, fig.height=20, fig.width=24}

# Graph of variables: default plot
fviz_pca_var(g.res.pca, col.var = "black", repel = TRUE,)

# Control variable colors using their contributions
fviz_pca_var(g.res.pca, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE) # Avoid text overlapping
             

# Variable contributions to the principal axes:
# Contributions of variables to PC1
fviz_contrib(g.res.pca, choice = "var", axes = 1, top = 20)
# Contributions of variables to PC2
fviz_contrib(g.res.pca, choice = "var", axes = 2, top = 20)
# Contributions of variables to PC3
fviz_contrib(g.res.pca, choice = "var", axes = 3, top = 20)

# select the top 15 contributing variables
cos2_12 <- fviz_cos2(g.res.pca, choice="var", axes = 1:2, top = 15)
cos2_12
cos2_12.df <- cos2_12$data
write.csv(cos2_12.df, "CLR_cos2_381.csv")
```

# Examining the spread of samples

Graphs and tables here are exploratory- not in the final manuscript

```{r, fig.height=12, fig.width=14}
# Extract the results for individuals
g.ind <- get_pca_ind(g.res.pca)
g.ind

# Coordinates of individuals
head(g.ind$coord)

# Graph of individuals
# 1. Use repel = TRUE to avoid overplotting
# 2. Control automatically the color of individuals using the cos2
    # cos2 = the quality of the individuals on the factor map
    # Use points only
# 3. Use gradient color
fviz_pca_ind(g.res.pca, col.ind = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )

quali.sup=382

fviz_pca_ind(g.res.pca, habillage=quali.sup, palette=ent_palette, col.var= "black", geom = "point", addEllipses = TRUE, ellipse.level=0.92, pointsize=3, ggtheme=theme_bw())
```



# Biplot of individuals and variables
```{r, fig.height=12, fig.width=14}

PCA <- fviz_pca_biplot(g.res.pca, habillage=quali.sup, palette=ent_palette, col.var= "black", repel = TRUE, geom = "point", addEllipses = TRUE, ellipse.level=0.92, pointsize=3, ggtheme=theme_bw()) 
PCA 
PCA_recorded <- recordPlot()


fviz_pca_biplot(g.res.pca, repel = TRUE, habillage=quali.sup, palette = ent_palette, geom = "point", select.var=list(cos2=0.4), col.var= "black", addEllipses = TRUE, ellipse.level=0.95, pointsize=3)
#ggsave("images/PCA.png")

PCA13 <- fviz_pca_biplot(g.res.pca, axes=c(1,3), habillage=quali.sup, palette=ent_palette, col.var= "black", repel = TRUE, geom = "point", select.var=list(cos2=0.4), addEllipses = TRUE, ellipse.level=0.92, pointsize=3, ggtheme=theme_bw())
PCA13

PCA23 <- fviz_pca_biplot(g.res.pca, axes=c(2,3), habillage=quali.sup, palette=ent_palette, col.var= "black", repel = TRUE, geom = "point", select.var=list(cos2=0.4), addEllipses = TRUE, ellipse.level=0.92, pointsize=3, ggtheme=theme_bw())
PCA23
```

# B) and C) Relative abundance plots for different enetrotypes

# B) 

```{r}
# top 25 genera
# select ps object
RA.ps <- readRDS("fig2-RA.ps.rds")

sample_data(RA.ps)$Enterotype
sample_data(RA.ps)$RA_Enterotype <- sample_data(RA.ps)$Enterotype


sample_data(RA.ps)$RA_Enterotype[sample_data(RA.ps)$RA_Enterotype == "PA2.1"]  <- "PA1"
sample_data(RA.ps)$RA_Enterotype[sample_data(RA.ps)$RA_Enterotype == "PA2.2"]  <- "PA2"
sample_data(RA.ps)$RA_Enterotype[sample_data(RA.ps)$RA_Enterotype == "PA2.3"]  <- "PA3"

sample_data(RA.ps)$RA_Enterotype

PA2.merge <- merge_samples(RA.ps, "RA_Enterotype")
sample_names(PA2.merge)

PA.merge.r <- transform_sample_counts(PA2.merge, function(OTU) OTU/sum(OTU))
```

# B) Relative abundance of phyla in different enterotypes 
```{r}

PA.merge.phy <- tax_glom(PA2.merge, taxrank="Phylum", NArm = FALSE)
PA.merge.phy.r <- transform_sample_counts(PA.merge.phy, function(OTU) OTU/sum(OTU))

library(RColorBrewer)
# Define the number of colors you want
nb.cols <- 14
mycolors12 <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)
# Create a ggplot with 18 colors 
# Use scale_fill_manual


#PA1
PA.merge.phy.top <- names(sort(taxa_sums(PA.merge.phy.r), decreasing=TRUE))[1:14]
PA.merge.top.phy.ps <- prune_taxa(PA.merge.phy.top, PA.merge.phy.r)
RA_phy <- plot_bar(PA.merge.top.phy.ps, fill = "Phylum") + guides(fill = guide_legend(nrow = 14)) +
  scale_fill_manual(values = mycolors12) + theme_bw() + theme(axis.text.x = element_text(angle = 0, hjust = 1))  + scale_y_continuous(breaks=seq(0,1,0.1))

RA_phy
ggsave("Fig2_RA_phy.png")
```

#c Relative abundance of genera in different enterotypes
```{r, fig.width=8, fig.height=7}
library(RColorBrewer)
# Define the number of colors you want
nb.cols <- 25
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)
# Create a ggplot with 18 colors 
# Use scale_fill_manual

#PA1
PA.merge.top <- names(sort(taxa_sums(PA.merge.r), decreasing=TRUE))[1:25]
PA1.merge.top.ps <- prune_taxa(PA.merge.top, PA.merge.r)
RA_genus <- plot_bar(PA1.merge.top.ps, fill = "Genus") + guides(fill = guide_legend(nrow = 25)) +
  scale_fill_manual(values = mycolors) + theme_bw() + theme(axis.text.x = element_text(angle = 0, hjust = 1))  + scale_y_continuous(breaks=seq(0,1,0.1))

RA_genus
#ggsave("images/RA_genus.png")
#p2
p2.abundance.df <- p2$data

```



```{r}
PCA_RA <- ggarrange(fviz_pca_biplot(g.res.pca, 
                          habillage=quali.sup, palette=ent_palette, col.var= "black", repel = TRUE, geom = "point", addEllipses = TRUE,
                          ellipse.level=0.92, pointsize=3, ggtheme=theme_bw()) +
  theme(plot.title = element_blank(), legend.text = element_text(size=12), legend.title = element_text(size=14)),  # First row with scatter plot
          ggarrange(plot_bar(PA.merge.top.phy.ps, fill = "Phylum") + guides(fill = guide_legend(nrow = 14)) +
  scale_fill_manual(values = mycolors12) + theme_bw() + theme(legend.text = element_text(size=12), legend.title = element_text(size=14)) + labs(x = "Enterotype") + theme(axis.text=element_text(size=12), axis.title=element_text(size=14)) + theme(axis.text.x = element_text(angle = 0, hjust = 1))  + scale_y_continuous(breaks=seq(0,1,0.1)),
                    plot_bar(PA1.merge.top.ps, fill = "Genus") + guides(fill = guide_legend(nrow = 25)) +  
  scale_fill_manual(values = mycolors) + theme_bw() + theme(legend.text = element_text(size=12), legend.title = element_text(size=14)) + labs(x = "Enterotype") +theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14)) + theme(axis.text.x = element_text(angle = 0, hjust = 1))  + scale_y_continuous(breaks=seq(0,1,0.1)),
          ncol = 2, labels = c("B", "C")), # Second row with box and dot plots
          nrow = 2, 
          labels = "A")                                        

ggsave("PCA_RA.png", plot=PCA_RA, scale = 1, width = 14, height = 16, units = "in", dpi = 300)

```

# D-G: alphadiversity comparison
```{r, Read in dataframe with alphadiversity info}
# base alphadiversity on an UNFILTERED, rarefied (reads=11111) dataset at ASV level. 

PA2.df <- read.csv("PA2_alphadiversity_compare.csv")
```



```{r, Add enterotype info, alphadiversity and otu abundance to ps object metadata}

my_comparisonsPA <- list(c("PA1", "PA2"), c("PA2", "PA3"), c("PA1", "PA3")) 
```

# D Observed

```{r, fig.width = 3.5, fig.height =3, alphadiversity}
# Perorm pairwise comparisons

ent_palette <- c("#F86B62", "#00BA38", "#619CFF")
compare_means(Observed ~ PA2,  data = PA2.df)

AObs <- ggboxplot(PA2.df, x = "PA2", y = "Observed", 
          fill = "PA2", palette = ent_palette,
          ylab = "Observed", xlab = "Enterotype") + guides(fill = "none") + 
  ggpubr::stat_compare_means(comparisons = my_comparisonsPA) +
ggpubr::stat_compare_means(label.y = 1500)   # Add pairwise comparisons p-value

# test for normality 
shapiro.test(PA2.df$Observed)
# Result shows not normal

AObs
#ggsave("images/AObs.png",
  #plot = last_plot(),
  #scale = 1,
  #width = 3.5,
  #height = 3,
  #units = c("in"),
  #dpi = 300)

kruskal.test(Observed ~ PA2, data = PA2.df)
pairwise.wilcox.test(PA2.df$Observed, PA2.df$PA2,
                 p.adjust.method = "BH")

pairwise.wilcox.test(PA2.df$Observed, PA2.df$PA2, p.adjust="bonferroni")
```

# E Chao1

```{r, fig.width = 3.5, fig.height =3, alphadiversity}
# Perorm pairwise comparisons
compare_means(Chao1 ~ PA2,  data = PA2.df)

AChao <- ggboxplot(PA2.df, x = "PA2", y = "Chao1", 
          fill = "PA2", palette = ent_palette,
          ylab = "Chao1", xlab = "Enterotype") + guides(fill = "none") + 
  ggpubr::stat_compare_means(comparisons = my_comparisonsPA) +
ggpubr::stat_compare_means(label.y = 1750)   # Add pairwise comparisons p-value

# test for normality 
shapiro.test(PA2.df$Chao1)
# Result shows not normal


kruskal.test(Chao1 ~ PA2, data = PA2.df)
pairwise.wilcox.test(PA2.df$Chao1, PA2.df$PA2,
                 p.adjust.method = "BH")


pairwise.wilcox.test(PA2.df$Chao1, PA2.df$PA2, p.adjust="bonferroni")

AChao
#ggsave("images/AChao.png",
  #plot = last_plot(),
  #scale = 1,
  #width = 3.5,
  #height = 3,
  #units = c("in"),
  #dpi = 300)

```

#F Shannon 

```{r, fig.width = 3.5, fig.height =3, alphadiversity}
# Perorm pairwise comparisons
compare_means(Shannon ~ PA2,  data = PA2.df)


AShan <- ggboxplot(PA2.df, x = "PA2", y = "Shannon", 
          fill = "PA2", palette = ent_palette,
          ylab = "Shannon", xlab = "Enterotype") + guides(fill = "none") + 
  ggpubr::stat_compare_means(comparisons = my_comparisonsPA) +
ggpubr::stat_compare_means(label.y = 7)   # Add pairwise comparisons p-value

# test for normality 
shapiro.test(PA2.df$Shannon)
# Result shows not normal


kruskal.test(Shannon ~ PA2, data = PA2.df)
pairwise.wilcox.test(PA2.df$Shannon, PA2.df$PA2,
                 p.adjust.method = "BH")


pairwise.wilcox.test(PA2.df$Shannon, PA2.df$PA2, p.adjust="bonferroni")

AShan

#ggsave("images/Shan.png",
 # plot = last_plot(),
  #scale = 1,
  #width = 3.5,
  #height = 3,
  #units = c("in"),
  #dpi = 300)
```

#G InvSimpson

```{r, fig.width = 3.5, fig.height =3, alphadiversity}
# Perorm pairwise comparisons
compare_means(InvSimpson ~ PA2,  data = PA2.df)


AInvSim <- ggboxplot(PA2.df, x = "PA2", y = "InvSimpson", 
          fill = "PA2", palette = ent_palette,
          ylab = "InvSimpson", xlab = "Enterotype") + guides(fill = "none") + 
  ggpubr::stat_compare_means(comparisons = my_comparisonsPA) + ggpubr::stat_compare_means(label.y = 190)   # Add pairwise comparisons p-value

# test for normality 
shapiro.test(PA2.df$InvSimpson)
# Result shows not normal


kruskal.test(InvSimpson ~ PA2, data = PA2.df)
pairwise.wilcox.test(PA2.df$InvSimpson, PA2.df$PA2,
                 p.adjust.method = "BH")


pairwise.wilcox.test(PA2.df$InvSimpson, PA2.df$PA2, p.adjust="bonferroni")

AInvSim

#ggsave("images/AInvSim.png",
  #plot = last_plot(),
  #scale = 1,
  #width = 3.5,
  #height = 3,
  #units = c("in"),
  #dpi = 300)

```

```{r, fig.width = 14, fig.height =3.2}

alphadiversity2 <- ggarrange(AObs, AChao, AShan, AInvSim, ncol = 4, labels = c("D", "E", "F", "G", widths = 3.5,
  heights = 3))
alphadiversity2

ggsave("alphadiversity.png", width = 14, height = 3.2,)
```
